Incentivizing Exploration with Heterogeneous Utilities

### 2018-02-02 algorithm.tex
wrote intro and preliminaries

\subsection{Assumptions}
In this section, we state several assumptions assumed by our analysis.  First define
\begin{align}
\Omega_{i,j}=\{\theta:B(\theta)=i, \hat{B}(\theta)=j\}, \nonumber 
\end{align}
which is the set of agent preferences whose best arm is arm $i$ and second best arm is arm $j$. With this definition, our analysis makes the following assumptions:

\begin{assumption} Let $F_{i,j}(y)$ be the marginal cumulative density function (or cumulative mass function if $F(\cdot)$ is a discrete distribution) of $(u_i-u_j)\cdot\theta$ conditioned on $\theta \in \Omega_{i,j}$. We assume $F_{i,j}(y)\leq My$ for all $y\in R^{+}$, $\forall i,j$. 
\label{A1}
\end{assumption}

@@

\begin{assumption} We assume $F$ has a compact support set. Without loss of generality, we assume $\theta\in [0,W]^m$.
\label{A2}
\end{assumption}

We use $R = \max_{\theta, i,j} \theta \cdot (u_i - u_j)$ to denote the maximum regret that can be incurred at each time.  Assumption~\ref{A2} shows that $R<\infty$.

\begin{assumption}
Denote $p=\min_{i}P(\{\theta: B(\theta)=i\})$. We assume $p>0$.
\label{A3}
\end{assumption}

Assumption~\ref{A3} means each arm $i$ has a strictly positive proportion of users for which that arm is best.

>>>>

\dkcomment{I moved the assumptions to an earlier section.
I am saving here some text that was in that subsection.}

We use $R = \max_{\theta, i,j} \theta \cdot (u_i - u_j)$ to denote the
maximum regret that can be incurred at each time.
Assumption~\ref{A2} implies that the maximum regret or payment
incurred in any one round is finite.

@@

### 2018-02-02 introduction.tex
wrote intro and preliminaries

Here we pursue an alternate direction to our analysis of incentivizing exploration with heterogeneous preferences: we consider approximation-style results comparing achievable algorithm performance across four different problem settings, described in the table below.

\begin{center}
\begin{tabular}{ c|c|c| } 
 \hline
     & Pulls arms directly & Needs to incentivize \\ 
\hline
 Perfect Info on Preferences & $C11$: The ``God'' Policy & $C12$ \\ 
 Partial Info on Preferences & $C21$ & $C22$: actual algorithm \\ 
 \hline
\end{tabular}
\end{center}

Columns indicate whether the principal pulls arms directly (left column) or incentivizes agents who in turn actually pull the arms. Rows indicate whether the principal knows the preferences of the agents (top row) or not (bottom row).   Problem settings correspond to cells and are described by the names C11, C12, and so on. An optimal algorithm for problem setting C11 has the best performance possible.

We wish to make two high-level points through this analysis:

\begin{itemize}
\item First, better performance is possible in setting C22 than in setting C21, in the sense we can find tight approximation guarantees relative to C11 that are better for C22 than for C21.  This is achieved by setting up incentives in C22 that still allow agents to select arms aligned with their preferences, providing selections of arms that contain more information than in C21.
\item Second, heterogeneity in preferences improves performance over what can be achieved in a setting with homogeneous preferences.  This would be in the spirit of David K.'s results that heterogeneity in preference for money can be exploited to reduce the budget required.
\end{itemize}

In this paper, we study incentivizing exploration with heterogeneous agent preferences.
In this problem,   
arms have unknown multivariate attributes, and agents have heterogeneous linear utility functions that map these attribute vectors onto utilities. Agents see noisy observations of attributes of arms pulled by all previous agents, and estimate an arms' attribute vector by the simple average of these observations.  Agents are selfish, and pull the arm with the largest estimated utility summed with an optional arm-specific incentivizing payment chosen by the principle.
We study strategies for choosing such incentive payments that seek to maximize the total utility derived by agents, subject to a limitation on the total incentive payment.  To accomplish this goal, a strategy must induce sufficient exploration to reveal arms' attributes, while still letting agents select myopically and according to their preferences sufficiently often that high-utility arms are chosen and incentive payments are kept small.

Our problem setting models online review aggregators like Amazon, Yelp, and Tripadvisor that host crowdsourced reviews.  Users of these websites wish to use the reviews hosted there to choose the product / restaurant / vacation (generically referred to as an ``item'') that is best according to their preferences.  These reviews provide not just cardinal feedback, but also a description of items' attributes that a user may consider together with their personal preferences to select their preferred item.  An item with few reviews might have inaccurate attribute estimates, leading users to avoid it even though it may actually be their best choice.  Without incentives, this situation may persist and decrease welfare for the platform's user base.  By offering incentives, either through price reductions by Amazon or coupons from Yelp or Tripadvisor, a platform may induce more reviews of unexplored items and provide more value over the long term.

Our problem setting also applies to crowd science platforms like eBird \cite{frazier2014incentivizing, sullivan2009ebird}. EBird guides birding enthusiasts through a website to explore and report their findings to the birding community. Each user report contains information about when, where and how they go birding and what birds they see and hear. EBird may wish to incentivize enthusiasts to explore less-explored birding locations and provide more accurate reports on these locations. Each enthusiast may have different preference over a location's attributes such as the diversity of bird species, weather, distance and safety. By offering enthusiasts incentivies to explore, eBird can create a more accurate body of reports and provide better value to the birding community.

% Something interesting to study would be to alter the search rankings to influence people
% Also which items to spam people about reviewing

In this problem context, we study a simple policy that usually exploits, incentivizing agents to pull an arm only when the set of agent utility functions that would pull this arm without incentives has probability below a time-varying threshold. In our paper, we assume all arms are some agents' best arm. Under this assumption, we prove that with $O(N^2)$ payment budget, this policy has $O(N^2+M(\log(T))^2)$ cumulative expected regret where $M$ is an upper bound on the limiting marginal probability density of agent utilities that are nearly indifferent between their best and the second best arm. If all agents' utility difference between their best and second best arm is bounded below by a positive number, which typically happens when the agent utility distribution is discrete, this policy achieve constant cumulative expected regret $O(N^2)$. The key difference between our problem setting and both the homogenous preference setting and the traditional multi-armed bandits setting is that we must incentivize agents to try suboptimal arms much less often, since all arms are some agents' best arm.  Essentially, heterogeneity provides free exploration.  These results suggest that heterogeneous agent preferences reduce but do not eliminate the need to incentive exploration, in relation to single-preference settings.

We broadly categorize the relevant previous literature into two categories based on whether there is money transfer.
With money transfer, \cite{frazier2014incentivizing} considers a problem setting where the principal pays agents money to explore. This work assumes all agents have equal value for money and provide a complete characterization of achievable reward with a fixed budget. \cite{han2015incentivizing} generalizes this framework to include agents with heterogeneous value for money, and to allow an external signal to provide partial information about this valuation. Under this setting, this work proves a bound on achievable reward as a function of the budget and the signal scheme.

Without money transfer but using information asymmetry, \cite{kremer2014implementing} considers a simple model where agents arrive to the principal one by one and there are only two actions at each time. \cite{mansour2015bayesian} generalizes \cite{kremer2014implementing} by allowing a finite number of actions at each time. \cite{mansour2016bayesian} considers a problem setting where there are multiple agents at each time and agents may interact with each other. In these papers, the principal provides each agent a recommendation at each time that is Bayesian incentive-compatible. They prove the principal can achieve constant regret when utilities are deterministic and logarithmic regret when utilities are stochastic.

We structure our paper as follows: Section~\ref{sec:prob} formulates our problem; Section~\ref{sec:ub} states our algorithm and proves that we can achieve $O(N^2+M(\log(T))^2)$ regret with $O(N^2)$ incentive budget; Section~\ref{sec:lb} constructs an example showing regret is $\Omega(\log(T))$ in the worst case, regardless of incentive budget.

>>>>

A large number of sites are designed to facilitate joint discovery,
sharing, and recommendations of content by large numbers of users.
Such sites include news, photo, or video sharing sites,
sites to review restaurants, hotels, or travel experiences,
online stores at which users write reviews (such as Amazon),
and citizen science projects
(such as eBird \citep{sullivan2009ebird} or Galaxy Zoo).
\dkcomment{Insert citations.}
By learning from the experiences of other users, individuals can make
better choices to improve their own experience.

Viewed more abstractly, the users are jointly exploring a space of
many options (products, news stories, photos, birdwatching sites,
patches of sky to train their telescopes on, $\ldots$),
with the implicit goal of identifying the ``best'' ones.
Therefore, such scenarios can be fruitfully modeled in a bandit
learning framework.
However, contrary to standard bandit settings, the utilities of the
decision makers (the users) are not aligned with the overall utility.
Societally (i.e., in a suitable aggregate over all users),
it would be desirable to engage in considerable exploration of
different options, so as to provide higher rewards for a large number
of future users.
However, individual users only interact with the site a limited number
of times, and therefore have little incentive for exploration.
A particular clean model is obtained when each user interacts with the
site only once, and hence has no intrinsic utility for exploration. 
This model has been the subject of prior work, and forms the basis of
the present submission.

In order to effect an outcome which is close to societally optimal,
it is necessary to provide exploration incentives to the individual users.
This was noted in two recent lines of work:
\citet{kremer2014implementing}
and \citet{mansour2015bayesian,mansour2016bayesian}
assume that the site (also called the \emph{principal}) has an
informational advantage in being the only one to observe the results
of past arm pulls.
(Such an assumption applies, for instance, to route recommendations in
driving.)
The principal can exploit this advantage and make recommendations to
the individual agents which it is in their best interest to follow.
\citet{frazier2014incentivizing} and 
\citet{han2015incentivizing} instead assume that the results of all
past arm pulls are publicly observable
(such as reviews on an online retail site).
They instead allow payments which the principal can offer to users as
a reward for pulling particular arms.
We follow the model of \citet{frazier2014incentivizing,han2015incentivizing}
and consider a multi-armed bandit model in which the principal can
offer payments to the users for pulling particular arms.

Past work has assumed that users are homogeneous, i.e., the expected
reward a user derives from an arm is the same for all users.%
\footnote{\citet{han2015incentivizing} assume that users are
heterogeneous in their tradeoff between utility derived from arm pulls
and utility derived from the principal's payment.}
In reality, users will have different preferences, e.g., gastronomic,
political, aesthetic, practical, etc.,
and the preferences can typically not be observed by the principal.
This heterogeneity presents both a challenge and an opportunity.
On the one hand, it necessitates the exploration of more arms,
so as to let all users discover their best option.
On the other hand, the heterogeneity also provides exploration
``for free,'' as users will not simply herd on one ``best'' arm.
\emph{Our goal is to explore the impact of user heterogeneity
on the tradeoff between societal utility and the principal's payments.}

At a high level, we model the setting as follows.
(Formal definitions are given in Section~\ref{sec:prob}.)
Arms and users (or \emph{agents}) are characterized by payoff-relevant
\emph{attribute} (or \emph{feature}) vectors,
both drawn from known distributions.
An agent's reward from pulling an arm is the inner product of their
vectors (plus noise).
When an arm is pulled, a noisy version of its attribute vector is
observed by everyone.
Agents are myopic and will pull the arm whose expected attribute
vector (based on past noisy observations) maximizes their reward.
The principal can incentivize agents to pull particular arms by
offering rewards for the specific arm.
The principal's goal is to keep the cumulative regret across all
agents small, while incurring only small total payments.
Our main theorm can be stated informally as follows:

\begin{theorem} \label{thm:main-intro}
Assume that for each arm, at least a constant fraction of the
population likes this arm best.
If furthermore, the density of ties in agent preferences between arms
vanishes (in a sense made precise in Section~\ref{sec:ub}),
there is a policy that achieves constant
expected regret $C$ and constant%
\footnote{The constants $C,C'$ depend on the number of arms and the
  smallest fraction preferring any one arm.} expected payment $C'$.
If near-ties have non-zero density,
then the regret is bounded by $O(C + C'' \log^2(T))$,
while the expected payment is still bounded by $C'$;
here, $C''$ depends on the ``tie density.''
\end{theorem}

The policy achieving the result of Theorem~\ref{thm:main-intro} is
quite simple: it mostly lets agents exploit arms, but incentivizes
them to explore when arms that appear close enough to being tied are
underexplored. It is presented in detail in Section~\ref{sec:ub}.
The result can be smoothly extended when there some arms that are not
preferred by any agents.

\dkcomment{I strongly dislike the ``Table of contents'' paragraph, and
  much prefer weaving pointers into the main text.}

### 2018-02-02 problem_setting.tex
wrote intro and preliminaries

\section{Problem Setting}
\label{sec:prob}

We have $N$ arms. Arm $i$ has a fixed but unknown attribute vector $u_i\in \mathbb{R}^{m}$. 
A stream of myopic selfish agents come to our system.  Agent $t$ has linear preferences over attributes described by a vector $\theta_t \in \mathbb{R}^m$ that is unknown to the principal and drawn i.i.d. from a known distribution $F(\cdot)$ with compact support. We refer synonomously to an agent and that agent's preference vector: when we say ``an agent $\theta$'', we mean ``an agent with preference vector $\theta$.''

Each agent $t$ chooses an arm to pull $A_t$, according to a process described below, and obtains utility $\theta_t \cdot u_{A_{t}}$.  The principal and all agents then see a noisy observation of the attribute vector of the pulled arm of the form $O_t=u_{A_{t}}+\epsilon_{t}$, where $\epsilon_t\sim N(0, \sigma^2 I_{m})$ is independent normally distributed noise, and $I_m$ denotes an $m$-dimensional identity matrix.  Although we assume a common variance across attributes for simplicity of presentation, our theoretical results hold if the variance differs.

At each time $t$, for each arm $i$, we (the principal) offer a non-negative payment $c_{i,t}\geq 0$ based on previous observations.
We assume that agent $t$ chooses to pull the arm that myopically maximizes the sum of this payment and an estimate of the utility obtained $\theta_t \cdot u_{i,t}$ where $u_{i,t}$ denotes the simple average of $O_s$ over all previous pulls of arm $i$. In this paper, we assume all arms have been pulled once at time $t=0$ and $u_{i,0}$ denotes a random draw from the arm attribute vector. For $t>0$, denote $u_{i,t} = \frac{\sum_{s<t} O_s 1\{A_s = i\} + u_{i,0}}{\sum_{s<t} 1\{A_s = i\}+1}$ and $A_t=\argmax_{i}\{\theta_t\cdot u_{i,t}+c_{i,t}\}$, breaking ties in favor of the arm with the highest incentive.  We use $c_t=c_{A_{t},t}$ to denote the actual incentive payment at time $t$.

This behavior may be recovered if agents are Bayesian and share a common non-informative prior distribution that is constant over $\mathbb{R}^m$ and know $\sigma^2$.  In this case, the posterior distribution on $u_{i}$ at time $t$ is multivariate normal with mean $u_{i,t}$, and the expected value of $\theta_t \cdot u_i$ under this posterior conditioned on $\theta_t$ is $\theta_t \cdot u_{i,t}$ (see equation $2.13$ in section $2.5$, \cite{Ge04}).  Alternatively, one may simply take our assumption that agents use the average as their estimate of an attribute value directly without such a Bayesian justification.

We define the regret at time $t$ as $r(t)=\max_{i}\{\theta_{t}\cdot u_{i}\}-\theta_t\cdot u_{A_t}$, and the cumulative regret up to time T as $R(T)=\sum_{t=1}^{T}r(t)$. Define the cumulative payment up to time T similarly as $C(t)=\sum_{t=1}^{T}c(t)$. 
As the principal, we want to find a strategy $\mathcal{A}$ under which both the cumulative expected regret $\mathbb{E}_{\mathcal{A}}[R(T)]$ and the cumulative expected payment $\mathbb{E}_{\mathcal{A}}[C(T)]$ are small.

To support later development, we define some additional notation.
We let $B(\theta)$ and $\hat{B}(\theta)$ refer to the index of the arm that is best and second best for an agent with preference vector $\theta$, $B(\theta) \in \argmax_i \theta \cdot u_i$ and $\hat{B}(\theta)=\argmax_{i\neq B(\theta)}\theta\cdot u_{i}$, breaking ties uniformly at random. We let $N(i,t)$ denote the number of pulls of arm $i$ at times up to and including $t$ plus $1$ (because of the initial pull), i.e. $N(i,t)=\sum_{s<t} 1\{A_s = i\}+1$.
We call time $t_{n}=\min_{t}\{\forall i, N(i,t)\geq n\}$ the {\it starting point of the $n^{th}$ round}. We call the set of times $[t_{n}, t_{n+1})$ the {\it $n^{th}$ round}.

>>>>

\section{Preliminaries}
\label{sec:prob}

We consider a MAB setting with \ARMNUM arms.
Arm payoffs are determined by $d$ \emph{attributes} or \emph{features};
hence, arms can be identified with vectors $\ArmV{i} \in \R^d$.
The \ArmV{i} are (adversarially) fixed, but unknown to the agents
or the principal.
Whenever arm $i$ is pulled, its current utility-relevant features are
determined as $\ArmV{i} + \NoiseV$, where \NoiseV is a mean-zero
Gaussian%
\footnote{For simplicity of notation, we assume that the variance
  $\sigma^2$ is uniform across time steps. This is not material for
  the analysis.}
noise vector $\NoiseV \sim \Normal{\vc{0}}{\sigma^2 I_{d}}$.
\dkcomment{Clarify if we can handle more general noise.}
Here, $I_d$ denotes the $d \times d$ identity matrix.

At each time $t$, a new user (or \emph{agent}) arrives,
whose feature vector (which we also call his%
\footnote{We use male pronouns to refer to users and female pronouns
  to refer to the principal. 
  \dkcomment{Feel free to switch if you prefer it the other way.}}
\emph{type}) $\AgentV{t} \in \R^d$ is drawn from a known distribution
\AgentDist.
Depending on context, we will identify a agents with their arrival time
$t$ or their type \AgentV{t}.
When agent $t$ pulls arm $\Pull{t} := i$,
he and all future agents observe a vector
$\ObsV{t} = \ArmV{i} + \NoiseV[t]$ for arm $i$,
and his reward is $\AgentV{t} \cdot \ObsV{t}$,
i.e., agents have linear preferences.

For each time $t$ and arm $i$, let \NumPull{t}{i} be the number of
times that arm $i$ has been pulled (strictly) before time $t$.
An agent at time $t$ estimates arm $i$'s attribute vector as the
average of vectors observed during past pulls of the arm:
$\ArmEV{t}{i} = \frac{1}{\NumPull{t}{i}+1} \cdot
(\ArmEV{0}{i} + \sum_{t'<t: \Pull{t'} = i} \ObsV{t'})$;
here, \ArmEV{i}{0} is a single draw $\ArmV{i} + \NoiseV$ for arm $i$.
(In other words, we assume that each arm is pulled once for free at time 0.)

Since each user only pulls an arm once, users are \emph{myopic}:
in the absence of incentives, user $t$ will pull an arm from
$\argmax_i \AgentV{t} \cdot \ArmEV{t}{i}$.
To incentivize users to explore more, the principal can offer
\emph{payments} $\Pay{t}{i}$ to user $t$ for pulling arm $i$.
Then, user $t$ will pull an\footnote{We assume that ties are broken in
  favor of an arm with largest payment \Pay{t}{i}; this assumption is
  only for notational convenience, and can of course be avoided by
  raising payments infinitesimally.} arm $i$ maximizing
$\argmax_i (\Pay{t}{i} + \AgentV{t} \cdot \ArmEV{t}{i})$.
The principal cannot observe \AgentV{t},
and only knows the distribution \AgentDist from which it is drawn.
Her goal is to achieve a good tradeoff between the cumulative
\emph{regret} experienced by all users up to time $T$,
and the total \emph{payment} she makes to the users.

We define the regret at time $t$ as
$\Regret{t} = (\max_{i} \AgentV{t} \cdot \ArmV{i}) - \AgentV{t} \cdot \ArmV{\Pull{t}}$,
and the cumulative regret up to time $T$ as
$\TotalRegret{T} = \sum_{t=1}^{T} \Regret{t}$.
Similarly, $\PayA{t} = \Pay{t}{\Pull{t}}$ is the actual incentive
payment at time $t$,
and the cumulative payment up to time $T$ is
$\TotalPay{T} = \sum_{t=1}^{T} \PayA{t}$.
More formally, the principal's goal is to find a policy
\POLICY for offering payments under which both the cumulative expected
regret
$\Expect[\POLICY]{\TotalRegret{T}}$ the cumulative expected payment
$\Expect[\POLICY]{\TotalPay{T}}$ are small.

To support the formulation of our results and the analysis,
we define the following additional notation.
We let
$\Best{\AgentV{}} \in \argmax_i \AgentV{} \cdot \ArmV{i}$
and
$ \Second{\AgentV{}} \in \argmax_{i \neq \Best{\AgentV{}}} \AgentV{} \cdot \ArmV{i}$
denote the (indices of) the best and second-best arms for an agent
with attribute vector \AgentV{};
\dkdelete{breaking ties uniformly at random}.
\dkcomment{What would it mean to break ties at random in a
  definition?}

\dkcomment{Do we need this definition here?}
We call time $t_{n}=\min_{t}\{\forall i, N(i,t)\geq n\}$ the
\emph{starting point of the $n^{th}$ round}.
We call the set of times $[t_{n}, t_{n+1})$ the
\emph{$n^{th}$ round}.

\dkdelete{This behavior may be recovered if agents are Bayesian and
  share a common non-informative prior distribution that is constant
  over $\mathbb{R}^m$ and know $\sigma^2$.  In this case, the
  posterior distribution on $u_{i}$ at time $t$ is multivariate normal
  with mean $u_{i,t}$, and the expected value of $\theta_t \cdot u_i$
  under this posterior conditioned on $\theta_t$ is $\theta_t \cdot
  u_{i,t}$ (see Equation 2.13 in Section 2.5, \cite{Ge04}).
  Alternatively, one may simply take our assumption that agents use
  the average as their estimate of an attribute value directly without
  such a Bayesian justification.}
\dkcomment{I wonder what's the best way to discuss this.}

\subsection{Properties of \AgentDist}
Our algorithms rely on a few assumptions about the agent distribution
\AgentDist.

\begin{assumption}[Compact Support] \label{A2}
\AgentDist has a compact support set contained in $[0,\Diam]^d$.
\end{assumption}

Let $\MinProb = \min_{i} \Prob[\AgentDist]{\Set{\AgentV{}}{\Best{\AgentV{}} = i}}$
denote the minimum (over all arms) fraction of users that prefer any
particular arm.

\begin{assumption}[Every arm is someone's best] \label{A3}
Each arm $i$ has a strictly positive proportion of users for whom $i$
is the best arm; that is, $\MinProb > 0$.
\end{assumption}

Let 
\begin{align*}
\FirstTwo{i}{i'}
  & = \Set{\AgentV{}}{\Best{\AgentV{}} = i, \Second{\AgentV{}} = i'}
\end{align*}
be the set of agent attribute vectors whose best arm is $i$ and
second-best arm is $i'$.
Let $F_{i,i'}$ be the marginal cumulative density function
(or cumulative mass function if $\AgentDist(\cdot)$ is a discrete distribution)
of $(\ArmV{i}-\ArmV{i'}) \cdot AgentV{}$,
conditioned on $\theta \in \Omega_{i,j}$.
In words, $F_{i,i'}$ is the distribution of the \emph{strength} of the
preference of a random agent for arm $i$ over arm $i'$, given that
$i$ and $i'$ are the agent's first and second choices.

\begin{assumption}[Not too many near-ties] \label{A1}
Near-ties have vanishing probability; 
that is, there exists a constant \TieDensity such that
for all pairs $i,i$ of arms and all $z \in \R^+$,
we have $F_{i,i'}(z) \leq \TieDensity \cdot z$.
\end{assumption}
